---
title: "Lab0"
format: html
editor: visual
embed-resources: true
---

## Demo

```{r, message = FALSE}
library("tidyverse")
library("palmerpenguins")
library("janitor")
library("readr")
glimpse(penguins_raw)
# penguins_raw |> glimpse()
# penguins_raw %>%  glimpse

old_names = colnames(penguins_raw)
penguins = penguins_raw %>% 
    janitor::clean_names()
new_names = colnames(penguins)
new_names
```

```{r, message = FALSE, warning=FALSE}
library("ggplot2")
penguins = penguins %>% 
    mutate(species = stringr::word(species, start = 1, end = 1))
penguins %>%  
  ggplot() + 
  # add the aesthetics
  aes(x = body_mass_g, 
      y = flipper_length_mm,
      colour = species) +
  # add a geometry
  geom_point() + 
  # tidy up the labels
  labs(x = "Body mass (g)",
       y = "Flipper length (mm)",
       colour = "Species")

#ggsave(filename = "myfirstggplot.png") to save the plot
```

## Exercises

1.  Generate a scatter plot for another pair of (numeric) variables

```{r, message = FALSE, warning=FALSE}
plt = penguins %>%  
  ggplot() +
  aes(x = culmen_length_mm, 
      y = culmen_depth_mm,
      colour = species) +
  geom_point() +
  labs(x = "Culmen Length (mm)", 
       y = "Culment Depth (mm)", 
       colour = "Species")
plt
```

2.  Colour by sex and use facet_wrap() to generate a plot for each species and island combination.

```{r, message = FALSE, warning=FALSE}
plt2 = penguins %>%  drop_na(sex) %>% 
  ggplot() +
  aes(x = culmen_length_mm, 
      y = culmen_depth_mm,
      colour = sex) +
  geom_point() +
  labs(x = "Culmen Length (mm)", 
       y = "Culment Depth (mm)", 
       colour = "sex")
plt2 + facet_wrap(vars(species, island))
```

3.  Try including a line of best fit by adding another geometry layer geom_smooth(method = "lm").

```{r, message = FALSE, warning=FALSE}
plt + geom_smooth(method = "lm")
```

4.  Use a different geometry, geom_histogram() to create a histogram for flipper length, coloured by species.

```{r plot, message = FALSE, warning= FALSE}
#| label: fig-figure
#| fig-cap: Palmer Penguins 

plt3 = penguins %>% 
  ggplot() +
  aes(x = flipper_length_mm, fill = species) +
  geom_histogram() + 
  labs(x = "Flipper Length (mm)", y = "Count", fill = "Species", title = "Flipper lengths of penguins by species")

plt3
```

5.  Save an updated version of your plot using ggsave().

This saves @fig-figure to the working directory.

```{r, message = FALSE, warning= FALSE}
ggsave(filename = "penguinsplot.png")
```

6.  Try outputting the data to a CSV file using the write_csv() function which can be found in the readr package.

```{r, message = FALSE, warning = FALSE}
write_csv(penguins, "penguins_csv")
```

## Advanced

```{r, message = FALSE, warning=FALSE}
# install.packages("plotly")
library("plotly")
myplot = penguins %>% 
  ggplot() +
  # add the aesthetics
  aes(x = body_mass_g,
      y = flipper_length_mm,
      colour = species) +
  # add a geometry
  geom_point() +
  # tidy up the labels
  labs(x = "Body mass (g)",
       y = "Flipper length (mm)",
       colour = "Species")
plotly::ggplotly(myplot)
```
