---
title: "Lab0"
output: 
  html_document:
    theme: cerulean
    toc: yes
    toc_float:
      collapsed: true
date: "2024-07-19"

---
# Demo
```{r, message = FALSE}
library("tidyverse")
library("palmerpenguins")
library("janitor")
glimpse(penguins_raw)

old_names = colnames(penguins_raw)
penguins = penguins_raw %>% 
    janitor::clean_names()
new_names = colnames(penguins)
new_names
```

```{r}
library("ggplot2")
penguins = penguins |>
    mutate(species = stringr::word(species, start = 1, end = 1))
penguins %>%  
  ggplot() + 
  # add the aesthetics
  aes(x = body_mass_g, 
      y = flipper_length_mm,
      colour = species) +
  # add a geometry
  geom_point() + 
  # tidy up the labels
  labs(x = "Body mass (g)",
       y = "Flipper length (mm)",
       colour = "Species")

#ggsave(filename = "myfirstggplot.png") to save the plot
```

# Exercises
1. Generate a scatter plot for another pair of (numeric) variables
```{r}
plt = penguins %>%  
  ggplot() +
  aes(x = culmen_length_mm, 
      y = culmen_depth_mm,
      colour = species) +
  geom_point() +
  labs(x = "Culmen Length (mm)", 
       y = "Culment Depth (mm)", 
       colour = "Species")
plt
```

2. Colour by sex and use facet_wrap() to generate a plot for each species and island combination.
```{r}
plt2 = penguins %>%  
  ggplot() +
  aes(x = culmen_length_mm, 
      y = culmen_depth_mm,
      colour = sex) +
  geom_point() +
  labs(x = "Culmen Length (mm)", 
       y = "Culment Depth (mm)", 
       colour = "sex")
plt2 + facet_wrap(vars(species, island))
```

3. Try including a line of best fit by adding another geometry layer geom_smooth(method = "lm").
```{r}
plt + geom_smooth(method = "lm")
```

4. Use a different geometry, geom_histogram() to create a histogram for flipper length, coloured by species.
```{r}
plt3 = penguins %>% 
  ggplot() +
  aes(x = flipper_length_mm, fill = species) +
  geom_histogram() + 
  labs(x = "Flipper Length (mm)", fill = "Species")

plt3
```

# Advanced
```{r, message = FALSE}
# install.packages("plotly")
library("plotly")
myplot = penguins |> 
  ggplot() +
  # add the aesthetics
  aes(x = body_mass_g,
      y = flipper_length_mm,
      colour = species) +
  # add a geometry
  geom_point() +
  # tidy up the labels
  labs(x = "Body mass (g)",
       y = "Flipper length (mm)",
       colour = "Species")
plotly::ggplotly(myplot)
```


