---
title: "Lab2"
format: 
  html:
    code-fold: true
    code-summary: "Show the code"
    embed-resources: true
toc: true
editor: visual
---

```{r, echo = FALSE, message = FALSE}
library(tidyverse)
```

# Exercises

## Dishonest Dice

A casino is worried about whether or not its die have been tampered with. To test this, a dealer rolls 4 dice 100 times and records how many even numbers (2, 4 or 6) appear.

```{=tex}
\begin{array} {lccccc}
\hline
\text{Number of evens} & 0 & 1 & 2 & 3 & 4 \\
\text{Number of rolls of 4 dice} & 1 & 15 & 42 & 32 &  10 \\ \hline
\end{array}
```
Can the scientist infer at the 5% significance level that the number of evens when $n = 4$ dice are rolled follows a binomial random variable with $p = \frac{1}{2}$?

Significance level: $\alpha = 0.5$.

Hypothesis:

$H_0$: The number of evens when 4 dice are rolled follows a binomial random variable $X \sim Bin(4, \frac{1}{2})$.

$H_1$: The number of evens when 4 dice are rolled does not follow this binomial distribution.

Assumptions:

The expected frequencies are sufficiently large (all greater than 5).

```{r}
y = c(1, 15, 42, 32, 10) # input the observed counts
x = 0:4 # define the corresponding groups
p = c(1, 4, 6, 4, 1)/16 # expected probabilities obeying the binomial distribution, could also use dbinom
n = 100
e = n*p
d = (y-e)^2 / e

df = tibble(
  User = c("0", "1", "2", "3", "4"),
  y = y,
  p = p,
  e = e,
  Contribution = d
)
df
```

Test statistic

```{r}
t = sum(d)
t
```

Using the chi-squared test with 4 degrees of freedom, we obtain a p-value of:

```{r}
p = pchisq(t, df = 4, lower.tail = F)
p
```

Conclusion:

Since the p-value is below the 5% significance level, we reject the null hypothesis in favour of the alternative hypothesis. Thus, the number of evens cannot be modelled via a random variable with $p = \frac{1}{2}$.

## Mammograms

Suppose that among 100,000 women with negative mammograms, 20 will have breast cancer diagnosed within 2 years; and among 100 women with positive mammograms, 10 will have breast cancer diagnosed within 2 years. Clinicians would like to know if there is a relationship between a positive or negative mammogram and developing breast cancer?

```{=tex}
\begin{array} {l|cc}
\text{Mammogram \ Breast cancer}& \text{Yes} &  \text{No}\\ \hline
\text{Positive} &      10 & 90\\  
\text{Negative} & 20 & 99,980 \\
\end{array}
```
```{r}
x = matrix(c(10, 20, 90, 99980), ncol = 2)
colnames(x) = c("Breast cancer: yes", "Breast cancer: no")
rownames(x) = c("Mammogram: positive", "Mammogram: negative")
```

1.  Is it appropriate to use a relative risk to quantify the relationship between the risk factor (Mammogram result) and disease (Breast cancer)? If so calculate the relative risk and provide an interpretation.

In this case, the sampling is performed on the risk factor (mammogram result). As a consequence, the relative risk is appropriate to quantify the relationship between risk factor and disease.

The relative risk:

```{r}
RR = (10/100)/(20/100000)
```

Women with a positive mammogram are 500 times more likely to develop breast cancer.

2.  Calculate the odds ratio of having breast cancer for positive vs negative mammograms and provide an interpretation.

Odds ratio

```{r}
OR = (10*99980)/(20*90)
```

The estimated odds of developing breast cancer for women with a positive mammogram are 555 times the estimated odds of developing breast cancer for women with a negative mammogram.

3.  Calculate a confidence interval for the odds-ratio, is there evidence that there might be a relationship between mammogram test results and breast cancer diagnosis?

Confidence interval for log-odds.

```{r}
SE = sqrt(1/10 + 1/100 + 1/20 + 1/100000)
logodds = log(OR)
L = logodds - 1.96*SE
R = logodds + 1.96*SE
ci = c(L, R)
ci
```

Confidence interval for odds ratio.

```{r}
exp(ci)
```

1 lies outside the confidence interval, therefore there is evidence that there might be a relationship between mammogram test results and breast cancer diagnosis.\

## Soccer Goals

Goals per soccer game arrive at random moments, and could be reasonably modelled by a Poisson process. If so, the total number of goals scored in a soccer game should be a Poisson random variable.

Here are the number of goals scored in each of the $n = 104$ games at the 2015 FIFA Women’s World Cup.

```{r}
goals <- c(
  1, 0, 0, 1, 1, 0, 0, 0, 1, 1, 2, 2, 4, 0, 10,
  0, 1, 1, 2, 3, 0, 4, 1, 3, 6, 0, 1, 0, 10, 1,
  2, 1, 0, 1, 1, 2, 3, 3, 3, 1, 2, 0, 0, 0, 0,
  1, 1, 1, 1, 1, 2, 0, 1, 0, 2, 2, 0, 1, 2, 1,
  1, 0, 1, 1, 0, 2, 2, 1, 0, 5, 2, 1, 4, 1, 1,
  0, 0, 1, 3, 0, 1, 0, 1, 2, 2, 0, 2, 1, 1, 1,
  0, 1, 0, 1, 2, 1, 2, 0, 2, 1, 0, 1, 5, 2
)
observed_goals = table(goals)
n = 104
observed_goals
```

To ensure expected frequencies are sufficiently large, collapse categories 4+.

```{r}
goal = c("0", "1", "2", "3", "4+")
fixed_goals = c(30, 40, 20, 6, 8)
names(fixed_goals) = goal
fixed_goals
```

Estimating $\lambda$ as the mean number of goals.

```{r}
parameter = sum(as.numeric(names(observed_goals)) * (observed_goals / 104))
```

Then

```{r}
k = c(0, 1, 2, 3)
probprelim = parameter^k * exp(-parameter) / factorial(k)
p = append(probprelim, 1-sum(probprelim))
p
e = 104 * p
e
d = (fixed_goals - e)^2 / e 

df = tibble(
  Goals = c("0", "1", "2", "3", "4+"),
  y = fixed_goals,
  p = p,
  e = e,
  Contribution = d
)
df
```

The test statistic would be:

```{r}
t = sum(d)
t
```

Using the chi-squared distribution with 3 degrees of freedom:

```{r}
p = pchisq(t, df = 3, lower.tail = F)
p
```

At the $\alpha = 0.5$ level of significance, we would therefore retain the null hypothesis. The number of soccer goals per game can thus be appropriately modelled using a Poisson random variable.
